Dynamics of thin liquid film flows studied with the lattice Boltzmann method

### 2023-07-24 chapters/Introduction.tex
Chapter two feedback Jens

\section{Thin film application}
\label{section:applications}
\begin{figure}
    \centering
    \includegraphics[width=0.75\textwidth]{graphics/41563_2003_Article_BFnmat788_Fig1_HTML.png}
    \caption{Film dewetting of a surface in the spinodial regime.
    Series (a) are experimental measurements with an AFM, while series (b) is generated using a numerical approach with the same parameters taken from~\cite{beckerComplexDewettingScenarios2003}.}
    \label{fig:becker_dewetting}
\end{figure}
Applications of the thin film equation cover a wide range of use cases. 
There is for example the problem of pesticide distribution in agriculture, a problem that requires that the pesticide sticks to the plant.
From a medical perspective, our lung and in particular the alveoli need to be covered with a thin film of biological surfactants.
Certain disease inhibit the surfactant production and therefore threaten partial or full collapse of the lung~\cite{doi:10.1146/annurev.fl.26.010194.002525}.
Coatings of surfaces is one of the problems that has been scientifically investigated for the last few decades.
There are various ways to coat a surface with arbitrary kinds of liquids.
What is common to all of them is the fact that the process heavily relies on the affinity between the to be coated surface and the coating agent~\cite{bonnWettingSpreading2009}.
If this affinity is too small the coating is unstable and therefore making the film prone to wrinkling~\cite{dasilvasobrinhoStudyDefectsUltrathin1999} or eventually even rupture~\cite{oronLongscaleEvolutionThin1997, crasterDynamicsStabilityThin2009, beckerComplexDewettingScenarios2003}, as shown in Fig~\ref{fig:becker_dewetting}.

\begin{figure}
    \centering
    \includegraphics[width=0.75\textwidth]{graphics/Coating_intro.pdf}
    \caption{(a) Schematics of the dip coating process. 
    A substrate is withdrawn from a fluid bath and due to capillarity and solid-liquid interactions some liquid adheres to the substrate.
    Most of the liquid however is due to viscous drag not drawn with the substrate.
    A small portion depending on the surface tension and atmospheric pressure is evaporating.
    (b) Schematics of a slot die coating process.
    A die is placed on top of a constantly moving foil.
    Through that die a liquid is pumped onto the moving foil.}
    \label{fig:dip_coating}
\end{figure}
For different object sizes but also for different materials there exists different coating approaches.
One of these approaches is the so called dip coating, as displayed in Fig.~\ref{fig:dip_coating}(a). 
Much like the name suggests the to be coated object is dipped into a reservoir of coating liquid~\cite{scriven_1988, darhuber2000selective, grosso2011exploit}.
While the process itself is straightforward the \textit{art} of dip coating is fairly complex.

From a theoretical point of view this problem has first been addressed by Landau and Levic~\cite{landau1988dragging} as well as Derjaguin~\cite{derjaguin1993thickness}.
Derived in the above work were two limits concerning the dynamics of the film during the extrusion.
In the low extrusion velocity limit, assuming the simple geometry of a plate, one can neglect the contribution due to gravity.
The resulting equation leads to a balance between capillary and viscous forces. 
The film that is drawn with the object is usually very thin and the resulting meniscus between the object and the fluid reservoir is in size proportional to the capillary length.
In fact the thickness of the coating film is roughly given as $h_{\infty} \propto l Ca^{2/3}$, where $l$ is the capillary length and $Ca$ the capillary number.
The capillary number $Ca$ is one among many non-dimensional numbers.
It compares viscous drag with surface tension, whereas the capillary length $l$ is a length scale that separates between gravity and capillary driven dynamics.
This can be understood simply by looking at droplets. 
As long as the volume is small the droplet will form a spherical cap.
If the volume becomes too large, a paddle will form with a distinct difference in shape from a spherical cap, because gravity can no longer be neglected.
In the converse limit of high extrusion velocity the capillary forces become subdominant.
Therefore viscous forces are being balanced with gravity. 
Using the same scaling argument as before the coating thickness can be approximated with $h_{\infty} \propto l Ca^{1/2}$.
Clearly these two regimes show that the extrusion velocity of the coated object has a significant impact, assuming that the surface tension and viscosity are constant parameters.

In case of die or slot die coating the to be coated object is not immersed into a reservoir of coating liquid.
Instead the coated material is usually an elastic foil which moves with constant velocity under a die.
A film is created because the coating fluid is pumped through the die which is placed on top of the foil with some distance $d$.
Similar to the illustration in Fig.~\ref{fig:dip_coating}(b), a film is drawn due to balance between capillary and viscous forces.
The thickness of the resulting film however does not depend on gravity in this case.
Ruschak showed that similar to the low extrusion velocity limit of the dip coating problem the film thickness after die coating can be approximated using $h_{\infty} = d~Ca^{2/3}$~\cite{RUSCHAK19761057}.

These two examples highlight an important behaviour.
The solutions consists of some variable, here the capillary number, that scales with a well defined exponent, e.g. $2/3$ and $1/2$.
These results can be addressed as dimensional analysis, a rather fancy word for getting the exponent correct because in the end the height or thickness needs to be of the dimension of a length.
They often support so-called self similar behaviour, thus upon rescaling results with a quantity it can be possible to derive a single master curve for all data.

\section{Literature Overview}
\label{section:literature}
\begin{figure}
    \centering
    \includegraphics[width=0.85\textwidth]{graphics/reynolds_slip_bearing.pdf}
    \caption{A drawing from Osborn Reynolds on slipper bearing lubrication.
    The four figures are taken from ref.~\cite{ReynoldsLubr} and cropped together in a single image.
    Displayed is the load balance between the lubricant and the solid body on top.}
    \label{fig:reynolds_work}
\end{figure}
Theory concerning lubrication is a little bit younger than the Navier-Stokes equation~\cite{Navier, Stokes}. 
Osborn Reynolds was the first to work on the problem of slipper bearing, as shown in Fig.~\ref{fig:reynolds_work}.
The idea of the slipper bearing is to reduce the wear of moving mechanical components due a lubrication fluid, in Reynolds case that fluid was olive oil~\cite{ReynoldsLubr}.
Within this work Reynolds developed what is called today the lubrication approximation.
Instead of solving the problem in all it's glory, Reynolds derived that due to constraints the problem can be simplified. 
The first constraint is the fact that the film's thickness $H$ is much smaller than it's lateral extension $L$ and therefore $\varepsilon = H/L \ll 1$. 
Besides this ratio the liquids inertia can be neglected as the acceleration of the fluid is slow.
Variations along the liquid interface are smooth thus the gradient $|\partial h/\partial x_i| \ll 1$ for both lateral dimensions.
This clean and novel deduction dates back already to the year 1886 and will be further discussed in Chap.~\ref{chapter:theory}.

Since the 19$^{\text{th}}$ century the theory of lubrication had gain much attention and got extended to account for more complex problems.
Over time it became it clear that problems such as wetting and capillarity can be studied within this framework~\cite{RevModPhys.57.827, oronLongscaleEvolutionThin1997, crasterDynamicsStabilityThin2009}.
To this end the term thin film equation was introduced, which is nothing more than the functional form that the lubrication approximation yields. 
Wetting for example does strongly depend on the intermolecular forces. 
The DLVO potential made it possible to account for these forces~\cite{derjaguin1941acta, verwey1947theory}. 
DLVO is an acronym from the name of the four scientists Derjaguin, Landau, Verwey and Overbeek who developed this theory. 
Initially this potential was used to understand the stability of colloidal suspensions, however for thin film theory it offers a quantification for the disjoining pressure~\cite{oronLongscaleEvolutionThin1997, Peschka9275, moulton_lega_2013, PhysRevE.63.011208}.
Pierre-Gilles de Gennes, who was awarded with the Nobel Prize in physics in 1991 for his work on soft matter physics and his students were ones of the first who understood that ``\textit{dirty}'' soft matter physics can help in understanding thin film problems~\cite{RevModPhys.57.827,doi:10.1021/la0116342,de2003capillarity}.

Of course, by end of the 20$^{\text{th}}$ century there were already some yet unexplained experimental findings associated to the thin film equation, e.g. the spectrum of capillary waves~\cite{doi:10.1021/ja01014a015}.
While the spectrum was quickly connected with thermal fluctuations it took theory some time to catch up.
From a theoretical point of view one has to account for the stochastic nature of thermal fluctuations.
Entry point for this problem is not the Navier-Stokes equation but the Landau-Lifshitz Navier-Stokes (LLNS) equation~\cite{Landau1987Fluid}.
The stochastic thin film equation is then, again, derived from the Landau-Lifshitz Navier-Stokes by integration along the vertical dimension.
This step however poses a hard problem, because the stochastic stress tensor in principle depends on the vertical dimension.
Grün and coworkers found an elegant way to circumvent the problem with a single multiplicative gaussian term~\cite{Grun2006, Mecke_2005, PhysRevLett.99.114503, PhysRevE.102.053105, PhysRevE.100.023108, PhysRevE.92.061002}, the idea will be revisited in Chap.~\ref{chapter:second_paper}.
Around the same time Davidovitch et al. came up with a similar form for the stochastic thin film equation, studying the spreading of a droplet~\cite{PhysRevLett.95.244505}.
The addition of a gaussian noise term made it therefore possible to reproduce the observed capillary wave spectrum and explain the difference in spreading behaviour.
A further impact of this term was the reduction of rupture times for spinodal dewetting thin films~\cite{Grun2006, PhysRevLett.99.114503}.
Becker et al. pointed out that the classical thin film equation comes short in prediction of the experimentally measured rupture times~\cite{beckerComplexDewettingScenarios2003}.
In fact the rupture time is overestimated by the classical thin film equation, adding however thermal fluctuations do decrease the rupture times~\cite{Duran_Olivencia2019, shah_van_steijn_kleijn_kreutzer_2019}.

Additions of complex dynamics on top of the partial differential equation that defines the thin film problems is not limited to thermal fluctuations.
On the one hand it is possible to add degrees of freedom to the fluid. 
For example one can study the effect of non-Newtonian shear behaviour~\cite{zhang2005non, myers2005application}. 
Or inclusion of arbitrary concentration fields to induce Marangoni-like flows~\cite{sultan_boudaoud_amar_2005, C5SM01603G, PhysRevLett.93.247803}.
Even the previous example of thermal fluctuations~\cite{Grun2006, Mecke_2005, PhysRevLett.95.244505, PhysRevE.104.034801}
On the other hand the substrate can be e.g. chemically or topographically patterned.
A modified substrate can lead to a very different understanding of wetting behaviour as shown in refs.~\cite{cassieWettabilityPorousSurfaces1944, WHYMAN2008355}.
In 2019 the Deutsche Forschungsgemeinschaft (DFG) funded the priority program 2171: ``\textit{Dynamic wetting of flexible, adaptive and switchable surfaces}''.
The goal of this program is to enlarge the understanding concerning the dynamics of thin liquid film on complex substrates.
Therefore substrates with internal degrees of freedom.
For example substrates with a grafted polymer brushes or special treatment to generate a self-cleaning super hydrophobic surface similar to the lotus leaf effect~\cite{doi:10.1021/acs.langmuir.0c03226, thiele2020gradient, doi:10.1021/acsnano.9b08211}.
Picking up the example of the lotus leaf, it is not necessary that the substrate is stiff.
In fact most organic structures are not as stiff as a silica substrate (or the glass of a window), therefore wetting of soft materials is as well considered in this priority program~\cite{doi:10.1146/annurev-fluid-010719-060147, https://doi.org/10.1002/nme.6567, chen2011short}.
More so with special surface treatment it is possible to control the substrates surface chemistry~\cite{xin2010reversibly, WANG200718}.
These stimuli respondent substrates are collected under the umbrella of the term \textit{switchable substrates}.
Allowing for the dynamic control of the substrates wettability with e.g. light sources~\cite{doi:10.1126/science.288.5471.1624, doi:10.1246/bcsj.20180076}.
\begin{figure}
    \centering
    \includegraphics[width=0.88\textwidth]{graphics/no_v_lig.pdf}
    \caption{(a) Equilibrium state of a dewetting thin film on a patterned substrate, i.e. droplets inside of the contact angle minima. 
    The color corresponds to the contact angle. 
    Yellow shows a high and blue a low contact angle, ranging from 10$^{\circ}$ to 30$^{\circ}$. 
    (b) Similar to the left panel a dewetting thin film on a ``switchable'' substrate. 
    Due to the function that controls the contact angle field liquid rivulets are formed.}
    \label{fig:morph_transition}
\end{figure}
This opens up a new avenue for microfluidic devices.
In Chap.~\ref{chapter:third_paper} we discuss the influence a switchable wettability can have on the dewetting of a thin film.
In fact the switching can induce morphological transitions as depict in Fig.~\ref{fig:morph_transition}.

@@

While it is easy to say that ``\textit{software should work}'' it is hard to guarantee.
The complexity of writing good source code and working applications is at least comparable to that of a scientific problem.
In fact, good software has become so complex that the workforce of a huge margin of employees of a company can be used up on a problem as ``simple'' as a search engine.
To make it possible to work in cooperation at this scale it is necessary to use so called \textit{best practises} and \textit{style guides}~\cite{sommerville2004software}.
Such style guides could for example define how to add and test new functions.

Scientific software is often seen as a secondary or by-product of research and sadly often threaded that way.
In contrast to the peer review process for scientific findings and publications, software that generated scientific results is rarely tested by reviewers.  
While there are high standards to industrial software, e.g. Ansys Fluent~\cite{matsson2020introduction}, these standards are quite low for the average scientifically developed source code.
This however is fairly troubling due two reasons.

Software or source code is prone to bugs because it is a creative process.
One approach to overcome the issues of bugs is the so called test driven development (TDD) approach~\cite{beck2003test}.
The simple idea is that, whenever a function is written it is accompanied with one or more test cases.
Using this principle creates a testing suite that (hopefully) ensures that functionality is not broken if changes to the source code are made.
For mature software projects, e.g. OpenFOAM~\cite{jasak2007openfoam, jasak2009openfoam, chen2014openfoam}, these suits can be even integrated into the continuous development/continuous integration (CD/CI) step and therefore be automated.
Often however when a student starts a new software project, functionality or performance is the main point of interest.
During the development step tests and even more important documentation are left aside.
Ultimately this does not only open the gate for various bugs but to some extent it also renders the software useless for other researchers.
Needless to say, my very own implementation of the solver, see Chap.~\ref{chapter:fourth_paper}, generated correct results due to the cancellation of two bugs.
Defining a test suite from the beginning of the project would have spared me valuable time and effort.

On the other hand, research became more than ever dependent on computational resources.
Clearly the main topic of this thesis, computational fluid dynamics is a compute heavy subdomain of fluid dynamics. 
However recent advances in artificial intelligence show a rising trend to their applicability for industrial and scientific problems~\cite{acemoglu20198, doi:10.1080/14685248.2020.1797059}
This advances comes at a price, leaving aside the discussion if artificial intelligence is of value at all.
One of the main pillars that generates trust in scientific results is their reproducibility.
Independent of the number of trials a stone, when flung up in the air, will fall to the ground following a well defined path\footnote{Given the stone is not flung with earth's escape velocity, $v\approx 11.2$km/s}.
Generating huge amounts of data without labels and explanation makes it by definition impossible to verify any given finding or result from that experiment.
Mislabeling data is not something new, it always happened from time to time with e.g. laboratory journals.
However exa-scaling computing pushes this problem to new severity.
Now or never is the time to install protocols. 
Guides on how to deal with scientific software such that data can be reproduced or even more important reused.
While this may not be of interest for most readers, to me it is an essential point.
That said, Chapter~\ref{chapter:fourth_paper} describes the software developed and used for this thesis. 
One can find the open source software repository that hosts the code, tests and documentation.
Upon pull or merge requests an automated testing suite performs tests to ensure no bugs break necessary functionality.
Furthermore by definition of version control it is possible to reproduce all results generated with that software.

>>>>

\section{Thin film application}
\label{section:applications}
\begin{figure}
    \centering
    \includegraphics[width=0.75\textwidth]{graphics/41563_2003_Article_BFnmat788_Fig1_HTML.png}
    \caption{Film dewetting of a surface in the spinodial regime.
    Series (a) are experimental measurements with an AFM, while series (b) is generated using a numerical approach with the same parameters taken from~\cite{beckerComplexDewettingScenarios2003}.}
    \label{fig:becker_dewetting}
\end{figure}
Applications of the thin film equation cover a wide range of use cases. 
There is for example the problem of pesticide distribution in agriculture, a problem that requires that the pesticide sticks to the plant.
From a medical perspective, our lung and in particular the alveoli need to be covered with a thin film with a well defined concentration of biological surfactants~\cite{hermansLungSurfactantsDifferent2015}.
Certain disease inhibit the surfactant production and therefore threaten partial or full collapse of the lung~\cite{doi:10.1146/annurev.fl.26.010194.002525}.

Coatings of surfaces is one of the problems that has been scientifically investigated for the last few decades.
There are various ways to coat a surface with arbitrary kinds of liquids.
What is common to all of them is the fact that the process heavily relies on the affinity between the to be coated surface and the coating agent~\cite{bonnWettingSpreading2009}.
If this affinity is too small the coating is unstable and therefore making the film prone to wrinkling~\cite{dasilvasobrinhoStudyDefectsUltrathin1999} or eventually even rupture~\cite{oronLongscaleEvolutionThin1997, crasterDynamicsStabilityThin2009, beckerComplexDewettingScenarios2003}, as shown in Fig~\ref{fig:becker_dewetting}.

\begin{figure}
    \centering
    \includegraphics[width=0.75\textwidth]{graphics/Coating_intro.pdf}
    \caption{(a) Schematics of the dip coating process. 
    A substrate is withdrawn from a fluid bath and due to capillarity and solid-liquid interactions (adhesion) some liquid adheres to the substrate.
    Balancing between adhesion and viscous drag most of the liquid is left in the bath.
    Depending on the surface tension and atmospheric pressure, evaporation can further thin the drawn film.
    (b) Schematics of a slot die coating process.
    A die is placed on top of a constantly moving foil.
    Through that die a liquid is pumped onto the moving foil.}
    \label{fig:dip_coating}
\end{figure}
For different length scales (m, cm, mm) and for different materials (metals, glasses, polymer compounds) there exists different coating approaches.
The industrial coating process of a five by five meter aluminium plate differs significantly from the process to coat a millimetric biological sensor.
Among the various coating approaches one is the so called dip coating, as displayed in Fig.~\ref{fig:dip_coating}(a). 
Much like the name suggests the to be coated object is dipped into a reservoir of coating liquid~\cite{scrivenPhysicsApplicationsDIP1988, darhuberSelectiveDipcoatingChemically2000, grossoHowExploitFull2011}.
While the process itself is straightforward the \textit{art} of dip coating is far from being trivial, meaning that this process requires tight tolerances on extraction velocity, material parameters and if possible a vibration free extraction.

From a theoretical point of view this problem has first been addressed by Landau and Levich ~\cite{landauDraggingLiquidMoving1988} as well as Derjaguin~\cite{derjaguinThicknessLiquidFilm1993}.
In their fundamental work they derived two limits concerning the dynamics of the film during the extrusion.
The low extrusion velocity limit assumes that gravity and as such inertia is small and can be neglected, at least for a simple plate geometry.
The resulting equation leads to a balance between capillary and viscous forces. 
The film that is drawn with the object is usually very thin and the resulting meniscus between the object and the fluid reservoir is in size proportional to the capillary length.
In fact the thickness of the coating film is roughly given as $h_{\infty} \propto l Ca^{2/3}$, where $l$ is the capillary length and $Ca$ the capillary number.
The capillary number $Ca$ is one among many non-dimensional numbers in fluid dynamics.
It compares viscous drag with surface tension, whereas the capillary length $l$ is a length scale that separates between gravity and capillary driven dynamics.
This can be understood simply by looking at droplets. 
As long as the volume is small the droplet will form a spherical cap.
If the volume becomes too large, a paddle will form with a distinct difference in shape from a spherical cap, because gravity can no longer be neglected.
In the converse limit of high extrusion velocity the capillary forces become subdominant.
Therefore viscous forces are being balanced with gravity. 
Using the same scaling argument as before the coating thickness can be approximated with $h_{\infty} \propto l Ca^{1/2}$.
Clearly these two regimes show that the extrusion velocity of the coated object has a significant impact, assuming that the surface tension and viscosity are constant parameters.

In case of die or slot die coating the to be coated object is not immersed into a reservoir of coating liquid.
Instead the coated material is usually an elastic foil which moves with constant velocity under a die.
A film is created because the coating fluid is pumped through the die which is placed on top of the foil with some distance $d$.
Similar to the illustration in Fig.~\ref{fig:dip_coating}(b), a film is drawn due to balance between capillary and viscous forces.
The thickness of the resulting film however does not depend on gravity in this case.
Ruschak showed that similar to the low extrusion velocity limit of the dip coating problem the film thickness after die coating can be approximated using $h_{\infty} = d~Ca^{2/3}$~\cite{ruschakLimitingFlowPremetered1976}.

These two examples highlight an important behaviour.
The solution consists of some variable, here the capillary number, that scales with a well defined exponent, e.g. $2/3$ in the low and $1/2$ in the high extrusion velocity scenario.
These results can be addressed as dimensional analysis, a rather fancy word for matching dimensions on both sides of the equation.
In the end the height or thickness needs to be of the dimension of a length.
They often support so-called self similar behaviour, meaning under change of one or more parameters it is possible to collapse data in a master curve. 
Another dynamic problem that exhibits self similar behaviour is droplet coalescence, which is shortly discussed in Chapter~\ref{chapter:fourth_paper} and liquid lens coalescence~\cite{hackSelfSimilarLiquidLens2020}. 
% I am here
\section{Literature Overview}
\label{section:literature}
\begin{figure}
    \centering
    \includegraphics[width=0.85\textwidth]{graphics/reynolds_slip_bearing.pdf}
    \caption{A drawing from Osborn Reynolds on slipper bearing lubrication.
    The four figures are taken from Ref.~\cite{reynoldsTheoryLubricationIts1886} and cropped together in a single image.
    In (a) surface CD is fixed while surface AB moves to the left with velocity U. 
    The fluid between the surfaces feels the force $F = \mu \frac{U}{h}$, with h being the separation distance.
    The velocity profile changes uniformly from U at AB to zero at CD. 
    Panel (b) displays the case when two parallel surfaces approach each other.
    The pressure acting on surface CD is shown on top of the surface.
    Each line between the surfaces shows the tangential forces and thus indicates the fluid velocity profile.
    Allowing surface AB to move to the left while surface CD approaches with the same pressure distribution as in (b) it is shown in (c).
    This case is a combination of (a) and (b) with non-uniform force distribution between the surfaces.
    The last panel (d) displays the case when the surfaces are not parallel.
    Surface CD is kept constant while surface AB is moved to the left.
    The curve on top displays the pressure the fluid exerts on surface CD.}
    \label{fig:reynolds_work}
\end{figure}
% Theory concerning lubrication is a little bit younger than the Navier-Stokes equation~\cite{Navier, Stokes}. 
Fluid dynamics in general is mathematically described by the Navier-Stokes equation~\cite{Navier, Stokes}.
Interestingly, this equation was not the starting point for the lubrication approximation.
Only in 1886, more than half a century after Navier's derivation, Osborn Reynolds started looking into a mechanical problem. 
He was the first to work on the dynamics of a slipper bearing, as shown in Fig.~\ref{fig:reynolds_work}.
The idea of the slipper bearing is to reduce the wear of moving mechanical components due to a lubricating fluid in Reynolds' case this fluid was olive oil~\cite{reynoldsTheoryLubricationIts1886}.
Within this work Reynolds developed what is called today the lubrication approximation.
Instead of solving the problem in all its glory, Reynolds derived that due to constraints the problem can be simplified. 
The first constraint is the fact that the film's thickness $H$ is much smaller than its lateral extension $L$ and therefore $\varepsilon = H/L \ll 1$. 
Besides this ratio the liquids inertia can be neglected as the acceleration of the fluid is slow.
Variations along the liquid interface are smooth and thus the gradient $|\partial h/\partial x_i| \ll 1$ for both lateral dimensions.
This clean and novel deduction dates back already to the year 1886 and will be further discussed in Chapter~\ref{chapter:theory}.

Since the 19$^{\text{th}}$ century the theory of lubrication has gained a lot of attention and got extended to account for more complex problems.
Over time, it became clear that problems such as wetting and capillarity can be studied within this framework~\cite{degennesWettingStaticsDynamics1985, oronLongscaleEvolutionThin1997, crasterDynamicsStabilityThin2009}.
To this end the term ``thin film equation'' was introduced, which is nothing more than taking the assumption of the lubrication approximation and removing the upper wall of Fig.~\ref{fig:reynolds_work}.
A thin film is therefore a liquid layer with a free liquid/vapour interface and the thin film equation is an evolution equation for a film thickness.
It is possible to derive one more model based on the lubrication approximation which is the viscous sheet model, as the name suggests this model describes a liquid layer with two free interfaces~\cite{savvaViscousSheetRetraction2009}.

By now it should be clear that contact line motion and thus wetting strongly depends on the affinity between liquid and solid. 
The DLVO potential is one potential which allows to account for these interaction forces~\cite{vActaPhysicochim1941, verweyTheoryStabilityLyophobic}. 
DLVO is an acronym from the names of the four scientists Derjaguin, Landau, Verwey and Overbeek who developed this theory. 
Based on the theory one can derive an interaction potential which than was used to understand the stability of colloidal suspensions, however for thin film theory it offers a quantification of the disjoining pressure~\cite{oronLongscaleEvolutionThin1997, peschkaSignaturesSlipDewetting2019, moultonEffectDisjoiningPressure2013, diezGlobalModelsMoving2000}.
Pierre-Gilles de Gennes, who was awarded with the Nobel Prize in physics in 1991 for his seminal work on soft matter physics and his students were ones of the first who understood that ``\textit{dirty}'' soft matter physics can be useful in understanding thin film problems~\cite{degennesWettingStaticsDynamics1985,degennesFluidWallSlippage2002,degennesCapillarityWettingPhenomena2004}.
The term dirty in the above sentence is an expression that de Genne used to distinguish soft matter systems from pure gas, liquid or solids. 
He and his students developed the theoretical tools to derive forces and potential and created the term ``soft interface''.
A soft interface is for example the surface of a thin film, because it is deformable and feels interaction with both the gas phase above and solid phase below~\cite{degennesCapillarityWettingPhenomena2004}.
De Genne not only found the soft matter field, but also made large contributions to our understanding of wetting, interface dynamics and among other things liquid crystals.

Of course, by the end of the 20$^{\text{th}}$ century there were already some yet unexplained experimental findings associated to the thin film equation, e.g. the spectrum of capillary waves~\cite{vrijRuptureThinLiquid1968}.
While the spectrum was quickly connected with thermal fluctuations it wasn't an easy challenge to find a consistent theoretical description.
From a theoretical point of view one has to account for the stochastic nature of thermal fluctuations.
The entry point for this problem is not the Navier-Stokes equation but the Landau-Lifshitz Navier-Stokes (LLNS) equation~\cite{landauFluidMechanicsLandau2013}.
While the mass flux is conserved at all times, heat transfer can violate Fouriers law at molecular scales~\cite{bellNumericalMethodsStochastic2007}.
Landau and Lifshitz derived a consistent stochastic stress term that accounts for these fluctuations and they added it to the Navier-Stokes equation.
Their stochastic term has zero mean, therefore the fluctuations do not create or destroy energy, and covariance proportional to thermal energy and viscosity. 
The stochastic thin film equation is then, again, derived from the LLNS equation by integration along the minor dimension, usually considered the vertical dimension.
This step however poses a hard problem, because the stochastic stress tensor in principle depends on the vertical dimension.
Grün and coworkers found an elegant way to circumvent the problem with a single multiplicative Gaussian term~\cite{grunThinFilmFlowInfluenced2006, meckeThermalFluctuationsThin2005, fetzerThermalNoiseInfluences2007, zhangNanoscaleThinfilmFlows2020, zhangMolecularSimulationThin2019, nesicFullyNonlinearDynamics2015}. 
This idea will be revisited in Chapter~\ref{chapter:second_paper}.
Around the same time Davidovitch et al., came up with a similar form for the stochastic thin film equation, studying the spreading of a droplet~\cite{davidovitchSpreadingViscousFluid2005}.
The addition of a Gaussian noise term made it therefore possible to reproduce the observed capillary wave spectrum and explain the difference in spreading behaviour.
A further impact of this term was the reduction of rupture times for spinodal dewetting thin films~\cite{grunThinFilmFlowInfluenced2006, fetzerThermalNoiseInfluences2007}.
Becker et al. pointed out that the classical thin film equation comes short in the prediction of the experimentally measured rupture times~\cite{beckerComplexDewettingScenarios2003}.
In fact the rupture time is overestimated by the classical thin film equation, adding however thermal fluctuations do lead to decreased rupture times~\cite{duran-olivenciaInstabilityRuptureFluctuations2019, shahThermalFluctuationsCapillary2019}.

Additions of complex dynamics on top of the partial differential equation that defines the thin film problems is not limited to thermal fluctuations.
On the one hand it is possible to add degrees of freedom to the fluid. 
For example one can study the effect of non-Newtonian shear behaviour~\cite{zhangNonNewtonianEffectsLubricant2005, myersApplicationNonNewtonianModels2005}. 
Non-Newtonian liquids have a different response to stress than e.g. water. 
A short discussion about shear thickening, thinning liquids, such as toothpaste, follows in Chapter~\ref{chapter:theory}.
Other than the liquids' rheology it is possible to include arbitrary concentration fields and introduce Marangoni-like flows~\cite{sultanEvaporationThinFilm2005, hermansLungSurfactantsDifferent2015, surSteadyProfileFingeringFlows2004}.
Thermal fluctuations pose another addition of complexity as shown above~\cite{grunThinFilmFlowInfluenced2006, meckeThermalFluctuationsThin2005, davidovitchSpreadingViscousFluid2005, zitzLatticeBoltzmannSimulations2021}.

On the other hand the substrate can be e.g. chemically or topographically patterned.
A modified substrate can lead to a very different understanding of wetting behaviour as shown in Refs.~\cite{cassieWettabilityPorousSurfaces1944, whymanRigorousDerivationYoung2008}.
Several research programs revolved around wetting, capillarity and the thin film equation. 
Within Europa the German Research Foundation (DFG) funded the priority program 1052: ``Wetting and Structure Formation'' from 1998 to 2004.
Karin Jacobs as project leader, one particularly interesting outcome was the fundamental work about dewetting morphology which was carried out by Becker et al.~\cite{beckerComplexDewettingScenarios2003}. 
Recently, in 2019, the DFG funded another priority program on wetting, called priority program 2171: ``\textit{Dynamic wetting of flexible, adaptive and switchable surfaces}''. 
The goal of this program is to enlarge the understanding concerning the dynamics of thin liquid films on complex substrates.
Complex substrates can for example be substrates with grafted polymer brushes~\cite{thieleGradientDynamicsModel2020}.
These brushes can swell somewhat similar to a sponge, allowing for special wetting and dewetting behaviour on such substrates. 
One treatment the substrate to generate a self-cleaning superhydrophobic surface similar to the lotus leaf~\cite{liAdaptationStyreneAcrylic2021, wongMicrodropletContaminantsWhen2020}.
Picking up the example of the lotus leaf, it is not necessary that the substrate is stiff.
In fact most organic structures are not as stiff as a silica substrate (or the glass of a window), therefore wetting of soft materials is as well considered in this priority program~\cite{andreottiStaticsDynamicsSoft2020, alandUnifiedNumericalModel2021, chenShortTimeWetting2011}.
More so with special surface treatment it is possible to control the substrates surface chemistry~\cite{xinReversiblySwitchableWettability2010, wangPhotoresponsiveSurfacesControllable2007}.
These stimuli responsive substrates are collected under the umbrella of the term \textit{switchable substrates}.
Allowing for the dynamic control of the substrates wettability with e.g. light sources~\cite{ichimuraLightDrivenMotionLiquids2000, sekiWideArrayPhotoinduced2018}.
\begin{figure}
    \centering
    \includegraphics[width=0.88\textwidth]{graphics/no_v_lig.pdf}
    \caption{(a) Equilibrium state of a dewetting thin film on a patterned substrate, i.e. droplets inside of the contact angle minima. 
    The color corresponds to the contact angle. 
    Yellow shows a high and blue a low contact angle, ranging from 10$^{\circ}$ to 30$^{\circ}$. 
    (b) Similar to the left panel a dewetting thin film on a ``switchable'' substrate. 
    Due to the function that controls the contact angle field liquid rivulets are formed.}
    \label{fig:morph_transition}
\end{figure}
This opens up a new avenue for microfluidic devices.
In Chapter~\ref{chapter:third_paper} we discuss the influence a switchable wettability can have on the dewetting of a thin film.
In fact the switching can induce morphological transitions as depict in Fig.~\ref{fig:morph_transition}.

@@

While it is easy to say that ``\textit{software should work}'' it is hard to guarantee.
The complexity of writing good source code and working applications is at least comparable to that of a scientific problem.
In fact, good software has become so complex that the workforce of a huge margin of employees of a company can be used up on a problem as ``simple'' as a search engine.
To make it possible to work in cooperation at this scale it is necessary to use so called \textit{best practises} and \textit{style guides}~\cite{sommervilleSoftwareEngineering2015}.
Such style guides could for example define how to add and test new functions.

Scientific software is often seen as a secondary or by-product of research and sadly often threaded that way.
In contrast to the peer review process for scientific findings and publications, software that generated scientific results is rarely tested by reviewers.  
While there are high standards to industrial software, e.g. Ansys Fluent~\cite{matssonIntroductionANSYSFluent2022}, these standards are quite low for the average scientifically developed source code.
This however is fairly troubling for two reasons.

Software or source code is prone to bugs because it is a creative process.
One approach to overcome the issues of bugs is the so called test driven development (TDD) approach~\cite{beckTestdrivenDevelopmentExample2003}.
The simple idea is that, whenever a function is written it is accompanied with one or more test cases.
To illustrate this idea, let's discuss an example written in Julia:
\begin{minted}{julia}
function add(a, b)
    return a + b
end

@test add(5, 3) == 8
\end{minted}
The function \textbf{add} returns the value of variable $a$ plus variable $b$.
Below the definition is a test, that passes if $5 + 3 = 8$ and fails if this not the case.
At first glance the test does not add any value, but at second thought one can imagine a scenario where $a$ and $b$ are complex numbers, or even worse where their type is not numerical.
The misuse of the simple function \textbf{add}, could in principle break a program.
Debugging of this situation may be straightforward, but with a test one knows exactly which function returns an error.
A more specific example to this thesis is collision operation in the shallow water lattice Boltzmann method.
On a two dimensional grid where each point has eight neighbours the function reads:
\begin{minted}{julia}
function BGKandStream!(state::LBM_state_2D, sys::SysConst)
    # All distribution functions
    fe0, fe1, fe2, fe3, fe4, fe5, fe6, fe7, fe8 = viewdists(state.feq)
    ft0, ft1, ft2, ft3, ft4, ft5, ft6, ft7, ft8 = viewdists(state.ftemp)
    fo0, fo1, fo2, fo3, fo4, fo5, fo6, fo7, fo8 = viewdists(state.fout)

    omeg = 1-1/sys.tau
